ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura G\'ati).}

### 2018-07-30 Latex/Draft_Summer2018/Summer2018.tex
more reading

In the previous section, we have documented results regarding the relation of current ICT investment and future productivity. Clearly, to rationalize the empirical responses of aggregate macroeconomic variables to ICT shocks, we need a two-sector approach that distinguishes ICT goods from other output. In this section, we present a simple two-sector model that rationalizes our structural VAR estimates.

% Need to add something here that says that we investigate whether a spillover can explain things.

The model presented has several common features with the two-sector environment of \cite{greenwood1997long} where one sector displays a faster technology than the other. Three main assumptions distinguish our model from the one by \cite{greenwood1997long}. First of all, we focus specifically on the ICT capital sector, rather than equipment, as the sector with faster technological improvement. This assumption will be necessary if we want to fit the fact that relative prices between ICT investment and the overall economy display a downward sloping trend.\footnote{In addition, we also rely on the results of \cite{whelan2003two} and \cite{oulton2007investment} to correctly define productivity and output share in our setting.} A second difference is that we consider two production functions - one for each sector. As \cite{oulton2007investment} shows, the two formulations are isomorphic, so our modeling choice has no implication for the resulting dynamics. The reason we choose this approach is because it allows for definitions of GDP and TFP in the model that are consistent with the way these variables are constructed in the national income accounts. Finally, we assume that the total ICT capital stock has a positive spillover effect on the production function of both sectors. The general-purpose nature of ICT capital works through its own diffusion because information sharing improves only when new communication technologies are widely spread over the population.

>>>>

In the previous section, we have documented results regarding the relation of current ICT investment and future productivity. Clearly, to rationalize the empirical responses of aggregate macroeconomic variables to ICT shocks, we need a two-sector approach that distinguishes ICT goods from other output. In this section, we present a simple two-sector model that rationalizes our structural VAR estimates. We also attempt to give an economic explanation for why a sector that accounts for a minuscule fraction of the entire economy can give rise to medium-run responses of the magnitude we documented in the previous section. As discussed below in more detail, we therefore entertain the possibility that there are spillover effects coming from the stock of ICT capital, and we confront the model with data to see if this explanation can be supported empirically.

The model presented has several common features with the two-sector environment of \cite{greenwood1997long} where one sector displays a faster technology than the other. Three main assumptions distinguish our model from the one by \cite{greenwood1997long}. First of all, we focus specifically on the ICT capital sector, rather than equipment, as the sector with faster technological improvement. This assumption will be necessary if we want to fit the fact that relative prices between ICT investment and the overall economy display a downward sloping trend. A second difference is that we consider two production functions - one for each sector. As \cite{oulton2007investment} shows, the two formulations are isomorphic, so our modeling choice has no implication for the resulting dynamics. The reason we choose this approach is because it allows for definitions of GDP and TFP in the model that are consistent with the way these variables are constructed in the national income accounts.\footnote{See also \cite{whelan2003two}.} Finally, and crucially, we assume that the total ICT capital stock has a positive spillover effect on the production function of both sectors.

What considerations lead us to augment both production functions with a spillover from ICT capital? First of all, the large, significant and persistent impulse responses of aggregates like TFP, GDP and consumption to an ICT shock are hard to square with the fact that the ICT sectoral share is less than 5\%. This seems to suggest that there is an additional, indirect effect coming from the ICT sector. A spillover is a natural way of capturing such an indirect effect.

Secondly, a long line of literature going back to the early 1990s postulates that ICT may be a general-purpose technology.\footnote{\cite{basu2003case}, \cite{basu2007information}, \cite{fernald2017disappointing} and \cite{oliner2000resurgence} and are some examples.} Following the classical definition of \cite{bresnahan1992general}, GPTs are technologies that alter the process of economy-wide production, and thus affect the productivity of the entire economy.

The general-purpose nature of ICT capital works through its own diffusion because information sharing improves only when new communication technologies are widely spread over the population.
